Towards Automated Refactoring of Code Clones in Object-Oriented Programming Languages\\- Work in Progress -

### 2019-06-30 sattose/paper.tex
Type  1R clones

\subsubsection{Type 1R clones} \label{sec:type1r}
Type 1 clones are identical clone fragments except for variations in whitespace and comments \cite{roy2007survey}. However, when two clone fragments are textually identical, it does not yet indicate that they are functionally identical. Although textually equal, method calls can still refer to different methods, type declarations can still refer to different types and even variables can be of a different type. In such cases refactoring opportunities could be invalidated. Therefore, we propose an alternative definition of type 1 clones. We have named it ``type 1R clones''. The ``R'' stands for refactoring-oriented (and may be less suitable for other analyses). Type 1R clones differ from type 1 clones as follows:
\begin{itemize}
  \item \textbf{Compare the equality of the fully qualified method signature for method references.} If an identifier is fully qualified, it means we specify the full location of its declaration (e.g. \texttt{com.simonbaars.fruitgame.Apple} for an \texttt{Apple} object). This way we can validate whether two method references, like method calls, are functionally equal. In the method signature, not only the fully qualified identifier of the method should be considered, but also the type of all its arguments. This way we can be sure that two potentially cloned method references do not point to overloaded variants (in a case that the data type of arguments is overloaded).
  \item \textbf{Compare the equality of the fully qualified identifier for type references.} This way we can be sure that two referenced types are actually equal, and that they are not just two types with the same name.
  \item \textbf{Compare the equality of the fully qualified identifier for variable usages.} Two cloned lines might use a variable with the same name, but different types. This might pose serious challenges on refactoring, as the variables might not concern the same object or primitive. To check this, we need to track the declaration of variables and from this infer the fully qualified identifier of its type.
  \item \textbf{Compare the equality of the fully qualified identifier for method references and call signatures.} This is to prevent a change in functionality after merging the clone, as methods with equal names/parameters can still contain different functionality.
\end{itemize}

>>>>

\subsubsection{Type 1R clones} \label{sec:type1r}
We propose an alternative definition of type 1 clones. This definition requires cloned fragments to be not just textually equal, but also functionally equal. Because of this, type 1R clones are a subset of type 1 clones. Although requiring fragments to be functionally equal, type 1R clones do not allow for change in implementation (like type 4 clones).

We check functional equality of two fragments by validating the equality of the fully qualified identifier (FQI) for referenced types, methods and variables. If an identifier is fully qualified, it means we specify the full location of its declaration (e.g. \texttt{com.sb.fruit.Apple} for an \texttt{Apple} object). This way we can validate whether two method references, like method calls, are functionally equal. This entails the following procedures to verify equality:

\begin{itemize}
  \item \textbf{Referenced Types:} For all types references in the code, we should compare their FQI with other code fragements to ensure functional equality.
  \item \textbf{Variables:} We trace the declaration of used variables to ensure that their type equals the type of another equally named variable.
  \item \textbf{Method Calls:} A codebase can have several methods with equal names, but a different implementation. Because of this, we compare the FQI of the method signature of called methods. A method signature includes the FQI for a methods' name, return value and each of its arguments.
\end{itemize}
